### 2017-10-20 daniel_thesis/toc.tex

\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\begin{romanpages}
	\tableofcontents
	%
	\phantomsection
	\addcontentsline{toc}{chapter}{List of Figures} % Add table index to the table of contents (toc).
	\listoffigures
	%
	\phantomsection
	\addcontentsline{toc}{chapter}{List of Tables} % Add table index to the table of contents (toc).
	\listoftables
	%
	\phantomsection
	\addcontentsline{toc}{chapter}{List of Algorithms} % Add table index to the table of contents (toc).
	\listofalgorithms
	%
	\printglossary[type=\acronymtype,title={Nomenclature}]
\end{romanpages}