### 2017-10-24 daniel_thesis/force_quad_triang_se.tex

\chapter{Analytical Forces Induced by Dislocations on Quadratic Triangular Surface Elements}
\label{c:quad_triang}
	\section{Forces Exerted by a Dislocation Line Segment on Quadratic Triangular Surface Elements}
	\label{s:f_quad_triang}
	\begin{figure}
		\centering
		\includegraphics[width=\linewidth]{force_calc_quad_triangle.pdf}
		\caption[Diagram of the analytical force calculation on quadratic triangular surface elements.]{Diagram of the line integral method used to find analytical expressions for the forces exerted by \kwdp{dl} on quadratic triangular \kwdp{se}.}
		\label{f:fqts}
	\end{figure}
\savearabiccounter