### 2017-10-24 daniel_thesis/preface.tex

\addtocounter{chapter}{-1}
\chapter{Preface}
\label{c:pre}
%
\section{Notation}
\label{s:nota}
%
For the sake of clarity the following notation conventions have been used:
\begin{itemize}
	\item \Kwdp{tensor} are denoted by sans serif bold italics, $ \tns{T} $.
	\item Matrices are denoted by serif bold roman, $ \mtx{M} $.
	\item Vectors are denoted by serif bold italics, $ \vec{V} $.
	\item Scalars are denoted by serif italics, $ S $.
	\item Host variables are denoted by $\hvar{a}$.
	\item Global device variables are denoted by $\dvar{a}$.
	\item Thread variables are denoted by $\tvar{a}$.
\end{itemize}
\section{Conventions}
All angles are given in radians unless stated otherwise. All pseudo-code is C-style and assuming C conventions (row-order, 0-start indexing, memory allocation).

Elements, be they \kwdp{se}, \kwdp{fe}, \kwdp{dls} or else, are denoted by $ e $ and their total as $ E $. Similarly nodes are denoted by $ n $ and their total $ N $. Parallel indices are denoted by $ \rvar{idx} $ to distinguish them from regular indices/counters $i,~j,~k$. For disambiguation purposes the exponential function will de denoted as $ \exp(x) $.
\section{Typesetting}
\label{s:typeset}
%
This document was typeset using a custom-made \LaTeX{} document class created by the author, compiled with \hologo{XeLaTeX}, and the bibliography was produced with \hologo{BibTeX}. The custom document class can be found in \href{https://github.com/dcelisgarza/latex_templssaates}{https://github.com/dcelisgarza/latex\_templates}.
\section{Diagrams}
\label{s:diag}
%
All diagrams are vector graphics drawn using the open-source image editing software \href{https://inkscape.org/en/}{InkScape} for its ease of use, elegant simplicity, high quality outputs, and ``Draw~Freely'' philosophy.
\savearabiccounter