### 2017-11-30 daniel_thesis/toc.tex
Figured out the mapping issue from the analytical force calculation to matlab nodal forces. Will proceed to implement mapping and split functions up.

\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\begin{romanpages}
	\tableofcontents
	%
%	\ifnum\value{figure}>0
	\phantomsection
	\addcontentsline{toc}{chapter}{List of Figures} % Add table index to the table of contents (toc).
	\listoffigures
%	\fi
	%
%	\ifnum\value{table}>0
	\phantomsection
	\addcontentsline{toc}{chapter}{List of Tables} % Add table index to the table of contents (toc).
	\listoftables
%	\fi
	%
%	\ifnum\value{algorithm}>0
	\phantomsection
	\addcontentsline{toc}{chapter}{List of Algorithms} % Add table index to the table of contents (toc).
	\listofalgorithms
%	\fi
	%
	\printglossary[type=\acronymtype,title={Abbreviations}]
\end{romanpages}